GSEA datasets for FDR benchmarking

### 2017-12-08 datasets/GSEA/gsea.Rmd
GSEA

library(EnsDb.Hsapiens.v86)
gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")
dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id,]
dsdObject <- DESeq(dsdObject)
res <- results( dsdObject, lfcThreshold=1 )

@@

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 15 ) %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=7)

@@

rejection_scatter_bins( sGSEA, "ind_covariate", threshold=15, bins=4, as_fraction=FALSE, supplementary=FALSE)
```

@@

getOverlapIndexes <- function( labs=c("lfdr", "ihw-a10"), alpha=0.1 ){
  rowSums( assays(sGSEA)[["qvalue"]][,labs] < alpha) == length(labs) &
  !(rowSums( assays(sGSEA)[["qvalue"]][,c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf")] < alpha, na.rm=TRUE ) > (length(labs)))
}

set1 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1 )
set2 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh"), alpha=0.1 )
set3 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf"), alpha=0.1 )

@@

dim(goRes)

goRes[set1,]

@@

>>>>

library(EnsDb.Hsapiens.v86)
gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")

dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id,]
dsdObject <- DESeq(dsdObject)
res <- results( dsdObject, independentFiltering=FALSE )

@@

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 15 ) %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=11)

@@

@@

getOverlapIndexes <- function( labs=c("lfdr", "ihw-a10"), alpha=0.1, sGSEA ){
  rowSums( assays(sGSEA)[["qvalue"]][,labs,drop=FALSE] < alpha) == length(labs) &
  !(rowSums( assays(sGSEA)[["qvalue"]][,c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf")] < alpha, na.rm=TRUE ) > (length(labs)))
}

set1 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1, sGSEA=sGSEA )
set2 <- getOverlapIndexes( c("bl-df03"), alpha=0.1, sGSEA=sGSEA )
set3 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh"), alpha=0.1, sGSEA=sGSEA )

@@

@@

# Same analysis but using the HSC-MPP dataset from Nina

dsdMouse <- "dsdHSCMPP.RData"
if( !file.exists( dsdMouse ) ){
  download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1", destfile="dsdHSCMPP.RData")
}
load(dsdMouse)
dseHSCMPP <- DESeq( dseHSCMPP)

res <- results( dseHSCMPP, contrast=c("conditions", "HSC", "MPP1"), independentFiltering=FALSE )
genes <- as.numeric( res$padj < 0.1 )
names( genes ) <- rownames( res )
genes[is.na( genes )] <- 0
table(genes)

### preparing gene length info and goSets ###
library("EnsDb.Mmusculus.v75")
txByGene <- transcriptsBy(EnsDb.Mmusculus.v75, "gene")
geneLength <- sapply( width( txByGene ), median )
geneLength <- geneLength[names(genes)]

file="goSetsMm.RData"
if(!file.exists(file)){
  stopifnot( require( biomaRt ) )
  mart <- useMart( "ensembl", "mmusculus_gene_ensembl" )
  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, filters="ensembl_gene_id", values=rownames(res))
  goSets <- goSets[!nchar( goSets$go_id ) == 0,]
  goSets <- with( goSets, split( go_id, ensembl_gene_id ) )
  save(goSets, file=file)
}else{
  load(file)
}

pwf <- nullp(genes, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )

res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )
sGSEAMouse <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(sGSEAMouse) <- "qvalue"
sGSEAMouse <- addDefaultMetrics( sGSEAMouse )

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 20)  %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=7)

rejections_scatter( sGSEAMouse, as_fraction=FALSE, supplementary=FALSE)

plotFDRMethodsOverlap( sGSEAMouse, supplementary=FALSE, order.by="freq")

set1 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf"), alpha=0.1, sGSEA=sGSEAMouse)
set2 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1, sGSEA=sGSEAMouse )
set3 <- getOverlapIndexes( c("bl-df03"), alpha=0.1, sGSEA=sGSEAMouse )

sizesDf <- data.frame( 
  size=c( rowData( sGSEAMouse )$ind_covariate[which(set1)], 
          rowData( sGSEAMouse )$ind_covariate[which(set2)], 
          rowData( sGSEAMouse )$ind_covariate[which(set3)] ),
  label=rep(c("set1", "set2", "set3"), c(sum(set1), sum(set2), sum(set3)) ) )
